Placing the spotted T Tauri star LkCa 4 on an HR diagram

### 2016-08-04 document/ms.tex
A few typo and miscellaneous corrections from a draft read through

Figures \ref{fig:Hband3x7} and \ref{fig:Kband3x7} in the Appendix show 42 of the $H-$ and $K-$band spectra on a log scale with a single random composite model spectrum overplotted.  The red and blue lines show the hot photosphere and cool spot spectra respectively.  We inspected the quality of the fits, generally finding remarkable agreement with most spectral orders.  Many orders demonstrated spectral line outliers, as noted in \iancze.  Most of the gross spectral features are captured by the composite model.  There are many examples of spectral lines that are present only in the hot photosphere model, or only in the cool photosphere model.  For example, the OH 1.563 $\mu$m line noted by \citet{oneal01} shows a clear pattern of 3 lines in the data, with the central line exceeding the depth of the adjacent two lines.

Best fit values for the stellar parameters summarized are listen in Table \ref{tbl_adopted_props}.  Remarkably, the filling factor of starspots exceeds 50\% with a best fit value of $c=75\pm 10 \% $.

@@

In \S 4.1 we describe how the two components of the photosphere fit low resolution spectra and optical/near-IR photometry.  The brightness, colors, spectral features should change as the spot rotates around the star.  In \S 4.2, we convert the $V$ brightness to a filling factor of the cool spot, and then in \S 4.2-4.3 we compare how the expected rotational modulation of photometry and spectral indices compares with observed measurements.

@@

Figures~\ref{fig:sed}-\ref{fig:lores} shows that the two-component photosphere matches the SED and flux-calibrated spectra well from $\sim 0.6-2.5$ $\mu$m.  The two components have fixed temperatures of 4200 K covering 30\% of the visible surface and 3000 K covering 70\% of the visible surface area.  The only free parameters are the luminosity and extinction, which are both scaled to match the spectrum.

The lower panels show photospheric spectra scaled separately to the observed red-optical and near-IR spectra.  The single temperature spectra are unable to reproduce the observed emission.  The $3900$ K component fits the TiO bands at reasonably well at $<7400$ \AA\ but fails to reproduce molecular bands at longer wavelengths.  The $3500$ K component hotter component suffers from the opposite problem, yielding molecular bands at short wavelengths that are too deep.  On the best-fit photosphere from \S 3 fits both the optical and the near-IR bands well.

Extinction paragraph, needs overhaul...  Relative to main sequence colors, the $E(B-V)=0.24$ would yield $A_V=0.72$ mag....  
The $J-K=0.93$ color is equal to the expected color to within 0.05 => 0.3 mag (for T=3000, 4200)... The $B-V$ is always dominated by the hot component and therefore does not depend on the filling factor of the hot component.  The $V-R$ varies more with $V$, indicating a larger contribution of the cool component.  The difference between observed and predicted $B-V$ color, $E(B-V)$, may correspond to the extinction.  After accounting for extinction synthetic model underpredicts both the $V-R$...

@@

Since 1992 the $V$ band brightness has varied from $12.3$ to $13.2$ mag, with a typical range of $0.5$ mag in any single epoch.  In a two-temperature atmosphere with fixed temperatures, this brightness corresponds directly to a filling factor for both components.  Figure~\ref{vband_spot} shows this conversion for temperatures of 4200 K, 3000 K, and a 75\% filling factor for the 3000 K component at $V_{est}=12.83$, as estimated at the time of the IGRINS spectrum.

The model photometry is obtained from the main sequence colors compiled by \citet{kenyon95}.  Colors from dwarfs stars should have minimal contribution from spots; measurements of pre-main sequence colors will include some unknown contribution of spots and cannot be used for this analysis.  Synthetic spectra of $3000-4000$ K stars predict $B-V$ colors that are much bluer than observed and that become bluer at cooler temperatures, the opposite of what is observed.  As a consequence, synthetic photometry also cannot be used here.

@@

The rotational modulation leads to much smaller brightness changes at near-IR wavelengths than at optical wavelengths (see Table~\ref{tab:photrange}).  The smaller amplitude of near-IR brightness is also seen in a few spotted stars in the optical/IR monitoring of NGC 2264 \citep{cody14}, although those stars have a much smaller amplitude than LkCa 4 in $V$ band.  Figure~\ref{fig:colors} demonstrates that the $B-V$ color from \citet{grankin08} does not depend on $V$, while the $V-R_J$ becomes slightly redder when $V$ is smaller.  The predicted colors from rotational modulation do not match the observed $V-R_J$ colors or the correlation with $V$, either because of changes in spot properties since the \citet{grankin08} data were obtained or because a two-temperature model for the photosphere is overly simplistic.

@@

The TiO band depths in these spectra are measured from spectral indices listed in Table~\ref{tab:tio}.  The TiO indices are also combined into a single index by calculating by  calculating how many standard deviations each point is from the median value of the index, and then averaging the standard deviation.

@@

Scaling our composite spectrum for LkCa 4 to the observed $K$-band photometry yields $R=2.2$ $R_\odot$ and $\log L/L_\odot=0.17$, assuming a distance of 131 pc based the parallax distance of the nearest Taurus member \citep{torres12}.   The spectrum of LkCa 4 is well reproduced by a cool component of $3000$ K covering 75\% of the surface and a $4200$ K component covering 25\% of the surface.  The effective temperature of these two components is 3611 K, as calculated from the average surface flux at the median $V$ band magnitude.  The apparent effective temperature in any single epoch would depend on the relative fractions of the hot and cool spot.  Decreasing the cool spot size to 60\% would increase the effective temperature to 3670 K, while increasing the spot size to 85\% would decrease the effective temperature to 3370 K.  {\bf GJH note:  holy crap!!!}.  Since the Hayashi tracks are nearly vertical at this temperature and age, the temperature uncertainty leads directly to an uncertainty in inferred mass.  Application of multi-temperature atmosphere is also likely to solve long-standing problems related to colors of young stars \citep[e.g.][]{stauffer03}

@@

{\bf paragraph MG-S:}  Comparisons to other results (both real and simulated) from single-temperature fits to APOGEE, GAIA-ESO wavelengths.  Approximate is ok...

@@

We also assumed that the spectral pattern of metallicity can be captured by a single coarse metric, $\Z$.  In reality, the low-temperature starspot spectrum of \name could exhibit $\A \neq 0$, resulting in a greater amplitude of residual spectrum correlations, and/or biased estimates for other parameters.  Our Gaussian process inference framework yields some resilience to the former problem.  For the latter, nonzero $\A$ would only impact a subset of all lines, and only to at a magnitude of typical departures of $\A$ in the solar neighborhood.  The large spectral grasp of IGRINS offers us some resilience to isolated orders badly affected by spectral line outliers.

@@

{\it Facilities:} \facility{Smith (IGRINS)}, \facility{AAVSO}, \facility{CFHT (ESPaDOnS)}, \facility{INTEGRAL (OMC)}, \facility{ASAS}, \facility{CrAO:1.25m}, \facility{ARC (TripleSpec)}, \facility{Hale (DBSP)}
{\it Software: } TBD
 \project{pandas} \citep{mckinney10}
    \project{emcee} \citep{foreman13},
 \project{matplotlib} \citep{hunter07},
 \project{numpy} \citep{vanderwalt11},
 \project{scipy} \citep{jones01},
 \project{ipython} \citep{perez07},
 \project{gatspy} \citep{JakeVanderplas2015},
 \project{starfish} \citep{czekala15}
 \project{seaborn} \citep{waskom14}
%\software{%
% \project{pandas} \citep{mckinney10}
%    \project{emcee} \citep{foreman13},
% \project{matplotlib} \citep{hunter07},
% \project{numpy} \citep{vanderwalt11},
% \project{scipy} \citep{jones01},
% \project{ipython} \citep{perez07},
% \project{gatspy} \citep{JakeVanderplas2015},
% \project{starfish} \citep{czekala15}}.

>>>>

Figures \ref{fig:Hband3x7} and \ref{fig:Kband3x7} in the Appendix show 42 of the $H-$ and $K-$band spectra on a log scale with a single random composite model spectrum overplotted.  The red and blue lines show the hot photosphere and cool spot spectra respectively.  We inspected the quality of the fits, generally finding good agreement with most spectral orders.  Many orders demonstrated spectral line outliers.  Most of the gross spectral features are captured by the composite model.  There are many examples of spectral lines that are present only in the hot photosphere model, or only in the cool photosphere model.  For example, the OH 1.563 $\mu$m line noted by \citet{oneal01} shows a clear pattern of 3 lines in the data, with the central line exceeding the depth of the adjacent two lines.

Best fit values for the stellar parameters summarized are listed in Table \ref{tbl_adopted_props}.  Remarkably, the filling factor of starspots exceeds 50\% with a best fit value of $c=75\pm 10 \% $.

@@

In \S 4.1 we describe how the two components of the photosphere fit low resolution spectra and optical/near-IR photometry.  The brightness, colors, and spectral features should change as the spot rotates around the star.  In \S 4.2, we convert the $V$ brightness to a filling factor of the cool spot, and then in \S 4.2-4.3 we compare how the expected rotational modulation of photometry and spectral indices compare with observed measurements.

@@

Figures~\ref{fig:sed}-\ref{fig:lores} show that the two-component photosphere matches the SED and flux-calibrated spectra well from $\sim 0.6-2.5$ $\mu$m.  The two components have fixed temperatures of 4200 K covering 30\% of the visible surface and 3000 K covering 70\% of the visible surface area.  The only free parameters are the luminosity and extinction, which are both scaled to match the spectrum.

The two panels in Figure \ref{fig:lores_twopanel} show photospheric spectra scaled separately to the observed red-optical and near-IR spectra.  The single temperature spectra are unable to reproduce the observed emission.  The $3900$ K component fits the TiO bands reasonably well at $<7400$ \AA\ but fails to reproduce molecular bands at longer wavelengths.  The $3500$ K component hotter component suffers from the opposite problem, yielding molecular bands at short wavelengths that are too deep.  On the best-fit photosphere from \S 3 fits both the optical and the near-IR bands well.

\todo{Extinction paragraph, needs overhaul...}  Relative to main sequence colors, the $E(B-V)=0.24$ would yield $A_V=0.72$ mag....  
The $J-K=0.93$ color is equal to the expected color to within 0.05 => 0.3 mag (for T=3000, 4200)... The $B-V$ is always dominated by the hot component and therefore does not depend on the filling factor of the hot component.  The $V-R$ varies more with $V$, indicating a larger contribution of the cool component.  The difference between observed and predicted $B-V$ color, $E(B-V)$, may correspond to the extinction.  After accounting for extinction synthetic model underpredicts both the $V-R$...

@@

Since 1992 the $V$ band brightness has varied from $12.3$ to $13.2$ mag, with a typical range of $0.5$ mag in any single epoch (\emph{c.f} Figure \ref{fig:PhotPhase}).  In a two-temperature atmosphere with fixed temperatures, this brightness corresponds directly to a filling factor for both components.  Figure~\ref{fig:vband_spot} shows this conversion for temperatures of 4200 K, 3000 K, and a 75\% filling factor for the 3000 K component at $V_{\rm est}=12.83$, as estimated at the time of the IGRINS spectrum.

The model photometry is obtained from the main sequence colors compiled by \citet{kenyon95}.  Colors from dwarf stars should have minimal contribution from spots; measurements of pre-main sequence colors will include some unknown contribution of spots and cannot be used for this analysis.  Synthetic spectra of $3000-4000$ K stars predict $B-V$ colors that are much bluer than observed and that become bluer at cooler temperatures, the opposite of what is observed.  As a consequence, synthetic photometry also cannot be used here.

@@

The rotational modulation leads to much smaller brightness changes at near-IR wavelengths than at optical wavelengths (see Table~\ref{tab:photrange}).  The smaller amplitude of near-IR brightness is also seen in a few spotted stars in the optical/IR monitoring of NGC 2264 \citep{cody14}, although those stars have a much smaller amplitude than LkCa 4 in $V$ band.  Figure~\ref{fig:colors} demonstrates that the $B-V$ color from \citet{grankin08} does not depend on $V$, while the $V-R_J$ becomes slightly redder when $V$ is smaller \comm{double check this statement}.  The predicted colors from rotational modulation do not match the observed $V-R_J$ colors or the correlation with $V$, either because of changes in spot properties since the \citet{grankin08} data were obtained or because a two-temperature model for the photosphere is overly simplistic. \comm{We could check for seasonal variations in V-R.}

@@

The TiO band depths in these spectra are measured from spectral indices listed in Table~\ref{tab:tio}.  The TiO indices are also combined into a single index by  calculating how many standard deviations each point is from the median value of the index, and then averaging the standard deviation.  \comm{this is confusing!}

@@

Scaling our composite spectrum for LkCa 4 to the observed $K$-band photometry yields $R=2.2$ $R_\odot$ and $\log L/L_\odot=0.17$, assuming a distance of 131 pc based the parallax distance of the nearest Taurus member \citep{torres12}.   The spectrum of LkCa 4 is well reproduced by a cool component of $3000$ K covering 75\% of the surface and a $4200$ K component covering 25\% of the surface.  The effective temperature of these two components is 3611 K, as calculated from the average surface flux at the median $V$ band magnitude.  The apparent effective temperature in any single epoch would depend on the relative fractions of the hot and cool spot.  Decreasing the cool spot size to 60\% would increase the effective temperature to 3670 K, while increasing the spot size to 85\% would decrease the effective temperature to 3370 K.  {\bf GJH note:  holy crap!!!}.  Since the Hayashi tracks are nearly vertical at this temperature and age, the temperature uncertainty leads directly to an uncertainty in inferred mass.

@@

{\bf paragraph MG-S:}  Comparisons to other results (both real and simulated) from single-temperature fits to APOGEE, GAIA-ESO wavelengths.  Approximate is ok... \todo{MGS provide text here}

@@

%Our Gaussian process inference framework yields some resilience to the former problem.

@@

{\it Facilities:} \facility{Smith (IGRINS)}, \facility{AAVSO}, \facility{CFHT (ESPaDOnS)}, \facility{INTEGRAL (OMC)}, \facility{ASAS}, \facility{CrAO:1.25m}, \facility{ARC (TripleSpec)}, \facility{Hale (DBSP)}

{\it Software: } 
 \project{pandas} \citep{mckinney10},
 \project{emcee} \citep{foreman13},
 \project{matplotlib} \citep{hunter07},
 \project{numpy} \citep{vanderwalt11},
 \project{scipy} \citep{jones01},
 \project{ipython} \citep{perez07},
 \project{gatspy} \citep{JakeVanderplas2015},
 \project{starfish} \citep{czekala15},
 \project{seaborn} \citep{waskom14}
%\software{%
% \project{pandas} \citep{mckinney10}
%    \project{emcee} \citep{foreman13},
% \project{matplotlib} \citep{hunter07},
% \project{numpy} \citep{vanderwalt11},
% \project{scipy} \citep{jones01},
% \project{ipython} \citep{perez07},
% \project{gatspy} \citep{JakeVanderplas2015},
% \project{starfish} \citep{czekala15}}.
